**Question 1**

Display the first six rows of the mpg dataset.

```{r}
library(ggplot2)
data(mpg)
head(mpg, n=6)
```


**Question 2**

Use ggplot to create a scatter plot with displ on the x-axis and hwy on the y-axis.


```{r}
ggplot(data=mpg, mapping=aes(x=displ,y=hwy))+
  geom_point()

#to add two different groups of data
ggplot(data=mpg, )+
  geom_point(aes(x=displ,y=hwy))+
  geom_point(aes(x=displ,y=cty,color='orange')) #define the color for the plot
  
       
```



**Question 3**

Modify the previous plot to color the points by class.

```{r}
ggplot(data=mpg, )+
  geom_point(aes(x=displ,y=hwy,color=class)) #class will color code the plot for you by class (defined in the dataset) #must be within aes() because it is variable and changes color based on the class

```



**Question 4**

Plot displ on the x-axis and cty on the y-axis. Use geom_point() to create the scatter plot.

```{r}
ggplot(data=mpg, )+
  geom_point(aes(x=displ,y=cty))
```



**Question 5**

Add a title to the last plot and label both x and y-axes.

```{r}
ggplot(data=mpg, )+
  geom_point(aes(x=displ,y=cty))+
  ggtitle("Vehicle Statistics")+
  xlab("Display")+
  ylab("City")
```


**Question 6**

Create a scatter plot with displ on the x-axis and hwy on the y-axis. Color the points by manufacturer.

```{r}
ggplot(data=mpg, )+
  geom_point(aes(x=displ,y=hwy,color=manufacturer))
```

**Question 7**

Create a box plot to visualize hwy miles per gallon across different class.

```{r}
ggplot(data=mpg,aes(x=class,y=hwy))+
  geom_boxplot()+
  labs(
    x="Class",
    y="MPG Highway",                   #can also label like this in ggplot
    title="Highway MGP by Vehicle Class"
  )
```


**Question 8**

Create a histogram of cty (city miles per gallon).

```{r}
ggplot(data=mpg, )+
  geom_histogram(aes(x=cty))
```


**Question 9**

Modify the histogram to have 30 bins.
```{r}
ggplot(data=mpg, )+
  geom_histogram(aes(x=cty), bins=30)
```



**Question 10**

Create a plot showing the density of hwy (highway miles per gallon).

```{r}
ggplot(mpg, aes(x = hwy)) +
  geom_density()
```



**Question 11**

Create a scatter plot with cty on the x-axis and hwy on the y-axis. Add a smoothing line with the method as "lm".

```{r}
ggplot(data=mpg, aes(x=cty,y=hwy))+
  geom_point()+
  geom_smooth(method=lm)
```



**Question 12**

Display the first six rows of the mpg dataset.

```{r}
head(mpg, n=6)
```



**Question 13**

Create a bar chart to count the number of cars in each class.

```{r}
class_freq <- as.data.frame(table(mpg$class))
class_freq

ggplot(data=class_freq, aes(x=Class,y=Freqency)+
  geom_bar(stat="identity")
  

```


**Question 14**

Create a bar chart to count the percentage of cars in each class. 

```{r}
class_freq$Percent <- (class_freq$Freq / sum(class_freq$Freq)) * 100
class_freq

colnames(class_freq) <- c("Class", "Frequency", "Percentage")
class_freq

```

