---
title: An R Markdown document converted from "in_class_exercices_for_loops.ipynb"
output: html_document
---

A 'for' loop in R is used to automate and repeat a task multiple times. It is a fundamental concept in programming, 
especially when you have a block of code that you want to execute repeatedly for different values.

Basic Structure:

for (item in list_of_items) {
    Inside this block, you can refer to 'item' to access the current element of 'list_of_items'
    do_something(item)
}

Here's a practical example:

Suppose you have a vector of numbers and you want to print each one squared:

```{r}
numbers <- c(1, 2, 3, 4, 5)
for (num in numbers) {
    squared <- num * num
    print(squared)
}
```

The above code will print the square of each number in the 'numbers' vector, one by one.

A couple of things to note:

The 'for' loop will iterate over each item in 'list_of_items' from the start to the end.
During each iteration, the loop assigns the current item to the variable 'item' and then executes the code inside the loop.
You can use the 'item' variable inside the loop to refer to the current item and perform operations on it.

### Question 1

Given a vector called `test_values` defined as follows:

```{r}
test_values <- c("Honolulu", "Los Angeles", "San Francisco", "New York")
```
Your task is to write a for loop that displays the position (or index) of each city in the vector followed by its name.

The output of your code should be

```bash
"1 Honolulu"
"2 Los Angeles"
"3 San Francisco"
"4 New York"
```

```{r}
for (i in c(1,2,3,4)){
   position <- (paste(i, test_values[i]))
    print(position)
}
```

### Question 2

You're given a vector named test_values:

```{r}
test_values <- c("New York", "Los Angeles", "San Francisco", "Honolulu")
```
Write a for loop to display the index of each city in the vector followed by a substring of the city's name. The substring should have a length equal to the index.

Expected output:

```bash
Copy code
1 N 
2 Lo
3 San
4 Hono
```

Hint: Use the substr() function to extract a substring. Here's how it works:

```{r}
city_name <- "Honolulu"
substr(city_name, start=1, stop=3)  # This will output "Hon"
```

```{r}
for (i in 1:4) {
    print(paste(i, (substr(test_values[i], start=1, stop=i))))
}
```

### Question 3
Without executing the code. What does the following code print?
```
my_data <- c(1,2,3,4,5)
for (my_val in my_data){
    print(my_data)
}
```

This will print:

1, 2, 3, 4, 5
1, 2, 3, 4, 5
1, 2, 3, 4, 5
1, 2, 3, 4, 5
1, 2, 3, 4, 5

It essentially prints then entire "my_data" vector for every value in my_data.

### Question 4

Given the vector values:

```{r}
values <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10)
```


Create a for loop to calculate the product of all the even numbers within this vector.


```{r}
accumulator <- 0
for (i in seq(2, 10, 2)) {
   accumulator <- values[i] + accumulator #take every other value in the sequence from position 2:10
   print(accumulator)
}
```

### Question 5
You have a vector of numbers:

```{r}
numbers <- c(1,2,3,4,5,6,7)
```

Create a for loop that multiplies each number with the previous ones and stores the result in a new vector called `products`.

Example:

For the first value in `products` (i.e., `products[1]`), the product is 1.
For the second value (i.e., `products[2]`) is 1*2 = 2.
For the third value (i.e., `products[3]`) is 1* 2* 3 = 6.
Now, write the for loop to fill the products vector with the results for each number in numbers.

```{r}
#for storing in an accumulator
products <- 1
for (i in 1:length(numbers)) {
    products <- products * numbers[i]
    print(products)
}
```
or
```{r}
#for storing in a vector
prod <- 1:6 #6 because 1 - length of the vector being used
for (i in 2:length(numbers)) { #2:7 because we want to start multiplication with the second value in numbers
    prod[i] <- numbers[i] * prod[i-1]
}
prod
```


