---
title: An R Markdown document converted from "in_class_practical_hypothesis_testing.ipynb"
output: html_document
---
```{r}
rm(list=ls())
### Exercise 1.
some_array <- 10:15
some_array
some_array[1] <- some_array[1] - 1
some_array[2] <- some_array[2] - 1
some_array[3] <- some_array[3] - 1
some_array[4] <- some_array[4] - 1
some_array[5] <- some_array[5] - 1
some_array

#loop
for(i in 1:5){
  print(i)
}
```

```{r}
#a <- rnorm(10, 0, 1)
#b <- rnorm(10, 0, 1)
#a
#b
#mean(a) - mean(b)

#create a list/array to store for loop in
list <- 1:10

#repeat this multiple times - use a for loop
for(batch in 1:10){
  a <- rnorm(10, 0, 1)
  b <- rnorm(10, 0, 1)
  print(mean(a) - mean(b))
  list[batch] <- (mean(a) - mean(b))
}

```


1. Generate 100 pairs of samples (s1 and s2) each of size 50 from the same distribution (mean = 10, std = 4) and plot the distribution of the difference between their means.
```{r}

```



2. Suppose you sampled two new samples and obtained a difference of 2.3, do you think the samples come from the same distribution?

```{r}


```



3. What is the p-value associated with such an event?



4. Now, upon sampling a mean of 0.2, what is the probability associated with such an event?




5. You have the following data which you obtained from an experiment consisting of <complete this>. Use the t-test to assess the difference between both samples. What does the t-test indicate about the data? Is it generated from the same distribution?



6. Would you be surprised if I told you that the data was in fact generated using the following method? Why?


7. What is the most appropriate observed statistic we should use to compare the two samples?



8. Plot the distribution of this test statistic, given the data above, under the null hypothesis.



9. What can you say about the test value obtained? Is it likely?



### Exercise 1.

Suppose a vendor selling a probiotic product claims that the microbial composition of their product is distributed among five species as follows:

Species_1: 30%
Species_2: 20%
Species_3: 20%
Species_4: 20%
Species_5: 10%

To investigate this claim, you decide to utilize Polymerase Chain Reaction (PCR) to quantify the proportions of Species 1-5 in a sample of their product. Upon analysis, you obtain the following proportions:

Species_1: 28%
Species_2: 21%
Species_3: 19%
Species_4: 23%
Species_5: 9%


1. What is a suitable statistic to compare the two datasets?

2. Write a function that takes two lists and returns the statistic you proposed above.

3. Generate the distribution of your statistic under the null hypothesis.

4. What is the p-value associated with observing such a test statistic under the null hypothesis?

5. Based on your statistical analysis, do the results agree with the manufacturer's claimed distribution of bacterial species? Explain your reasoning.

### Question 3

* Suppose we have the following two samples. Use a permutation-based test to assess whether these two samples come from the same or different distributions. Illustrate all the steps.

```{r}
sample_1 <- c(2.28, 2.55, 5.4, 3.01, 4.63, 1.56, 2.81, 3.38, 3.3, 1.99, 3.02, 1.92, 3.71, 4.08, 0.78, 4.24, 1.76, 3.45, 3.66, 2.8, 2.35, 3.17, 3.44, 3.88, 0.95, 1.36, 4.43, 4.05, 3.44, 3.72, 3.92, 0.34, 4.11, 2.52, 3.23, 2.7, 3.87)
sample_2 <- c(2.88, 3.92, 2.83, 1.99, 4.75, 4.19, 5.37, 3.38, 4.65, 5.67, 2.96, 1.71, 3.01, 3.04, 3.48, 5.35, 2.91, 3.75, 3.03, 3.32, 3.68)
```


