### Exercise 1

Generate a plot of the differences in means between 100 pairs of samples. Each sample pair (s1 and s2) should have 50 observations drawn from the same distribution (mean = 10, standard deviation = 3).

To assist you, I've provided a function named generate_diff_means. Here's how to use it:

Function Parameters:

* n: Number of sample pairs to generate. 
* size_1: Size of the first sample in each pair.
* size_2: Size of the second sample in each pair.
* mu_1: Mean of the first sample.
* sd_1: Standard deviation of the first sample.
* mu_2: Mean of the second sample.
* sd_2: Standard deviation of the second sample.

Output:
The function returns $n$ values, each representing the difference between the means of the two samples.

You can use this function just like the others we've encountered. It's absolutely okay if you're not yet familiar with how it works.

```{r}
generate_diff_means <- function(n, size_1, size_2, mu_1=0, sd_1=1, mu_2=0, sd_2=1){
    diff_means = 1:n
    for (sample_number in 1: n){
        s1 = rnorm(size_1, mu_1, sd_1)
        s2 = rnorm(size_2, mu_2, sd_2)
        diff_means[sample_number] <- mean(s1) - mean(s2)
    }
    diff_means
}
```

```{r}
# Parameter definitions
library(ggplot2)

n <- 100
size_1 <- 50
size_2 <- 50
mu_1 <- 10
sd_1 <- 3
mu_2 <- 10
sd_2 <- 3

#creating the function
generate_diff_means <- function(n, size_1, size_2, mu_1=0, sd_1=1, mu_2=0, sd_2=1){
    diff_means = 1:n
    for (sample_number in 1: n){
        s1 <- rnorm(size_1, mu_1, sd_1)
        s2 <- rnorm(size_2, mu_2, sd_2)
        diff_means[sample_number] <- mean(s1) - mean(s2)
    }
    diff_means
}

#fill in the parameters for the function created - applying the fuction to get actual values
mean_diff <- generate_diff_means(n=100, size_1=50, size_2=50, mu_1=10, sd_1=3, mu_2=10, sd_2=3)
mean_diff
```


2. Suppose you sampled two new samples and obtained a difference of 2.3, do you think the samples come from the same distribution?

```{r}
# Write your answer here
#These new samples are most likely to be from a difference distribution.
```

3. What is the empirical  p-value associated with such an event?

```{r}
# Write your code here
sum(abs(mean_diff) >= 2.3) / length(mean_diff)

```

4. Now, upon sampling a mean of 0.2, what is the empirical probability associated with such an event?

```{r}
# Write your code here
sum(abs(mean_diff) >= 0.2) / length(mean_diff)
```

5. You have the following two datasets, which you obtained from an experiment. Use the t-test (t.test function in R) to assess the difference between both samples. What does the t-test indicate about the data? Is it generated from the same distribution?

``` {r}
dataset_1 <- c( 9.22, 11.12, 8.91, 10.9, 9.91, 9.74, 9.39, 10.77, 10.51, 8.78, 8.86, 10.15, 10.4, 8.89, 9.82, 10.89, 10.26, 10.72, 8.72, 9.03, 9.99, 7.65, 9.7, 9.2, 10.86 )
dataset_2 <- c( 12.21, 10.14, 10.84, 9.31, 9.8, 10.16, 9.25, 11, 10.09, 9.34, 9.02, 9.84, 9.99, 10.98, 11.16, 12.02, 11.17, 9.87, 10.32, 9, 9.8, 10.38, 11.06, 10.65, 10.1 )
```

```{r}
# Write your code here
t.test(dataset_1, dataset_2)
```


```{r}
#These datasets are significantly different with a p-value of 0.03892. They were generated using a normal (Gaussian) because a t-test requires them to both be the same, but they were generated y=using different parameters according to the t-test.
```

6. Would you be surprised  if I told you that the data was in fact both were generated using the following method? Do you think it's possible? Why?

```
rnorm(25, 10, 1)
```

```{r}
#I would be be surprised because the t-test showed significant difference between the two data sets. The parameters could include the same sample size and mean, but the second data set might have a larger standard deviation of 2.

```


7. What is the most appropriate observed statistic we should use to compare the two samples using permutation?

```{r}
#You would use the difference of means as your observed statistic. The equivalent of running a non-parametric t-test is a permutation. 
```

Create a function named diff_between_means that calculates and returns the difference between the means of two given samples, `s1` and `s2`.

Function Usage:
```
diff_between_means(s1, s2)
```

for example 
```
diff_between_means(s1=c(1, 2, 3), s2=c(1, 2, 3))
```

the function should return 0. This is because both samples have the same mean, so the difference is 0.
and 
```
diff_between_means(s1=c(3, 3, 3), s2=c(1, 1, 1))
```
the function should return 2. The difference between their means is 3 - 1 = 2.

```{r}
# Write your function here
diff_between_means <- function(s1, s2){
  mean(s1) - mean(s2)
}
#testing the function according to parameters given above, should return 0
diff_between_means(s1=c(1, 2, 3), s2=c(1, 2, 3))
#testing the function according to parameters given above, should return 2
diff_between_means(s1=c(3, 3, 3), s2=c(1, 1, 1))
```

Remember our discussion about the importance of testing functions. I emphasized the use of assertions to make sure the functions we implement do what they are intended to do. For your convenience, I've written some assertions for the `diff_between_means` function. To test your function, you need to first install the `testit` package if you haven't already, 

install.packages("testit")

After defining your diff_between_means function, execute the provided assertions to verfiy if function works as expected.

```{r}
library(testit)
#if it doesnt like it, it will complain and return an error.
assert(
    diff_between_means(s1=c(1, 2, 3), s2=c(1, 2, 3)) == 0     
)
assert(
    diff_between_means(s1=c(3, 3, 3), s2=c(1, 1, 1))  == 2    
)
# IT
```

In class, we learned about a method called the permutation test. Here's a breakdown of what it does:

1. Combine the Datasets:
   If:
   s1 <- c(1, 2, 3, 4)
   s2 <- c(5, 6)
   Then, combined:
   s <- c(1, 2, 3, 4, 5, 6)

2. Shuffle the combined set, so s_shuffled  may become c(6,1,4,5,3,2)

3. Split the Shuffled Set:
   Using our shuffled list, we are required to split in two sets of sizes equal to s1 and s2. 
   s1_new becomes c(6,1,4,5)
   s2_new becomes c(3,2)

4. Calculate the Difference:
    mean(s1_new) - mean(s2_new) will be 2.5

Repeat these steps multiple times (e.g., 1000) to get a list of all differences.

I am providing a function called `bootstrap_null`, which automates the above steps. Since you alredy wrote a function to accomplish step 4, `bootstrap_null` will use the function you wrote earlier, showcasing function reusability one of the many advantages of using functions.

You us `bootstrap_null` as follows
```
bootstrap_null(n, s1, s2)
```
Where:
- n: Number of repetitions
- s1: First sample
- s2: Second sample

```{r}
bootstrap_null <- function(n, s1, s2){
    diff_means <-  1:n
    for (i in 1:n){
        s <- c(s1, s2)
        s_shuffled <- sample(s)
        s1_new <- s_shuffled[1:length(s1)]
        s2_new <- s_shuffled[(length(s1) + 1):length(s)]
        diff_means[i] <-  mean(s1_new) - mean(s2_new)
    }
    diff_means
}
```

9. Execute the function using `n = 1,000` and the `dataset_1` and `dataset_2` from question 5. Store the results in a variable called `perm_diff_means`

```{r}
# Write you code here
perm_diff_means <- bootstrap_null(n=1000, s1=dataset_1, s2=dataset_1)
```

```{r}
# this is an assert to make sure the function does indeed compute 10k means

assert( 
    length(perm_diff_means) == 1000
)    
```

10. Visualize the distribution of mean differences obtained from the permutation test, as recorded in `perm_diff_means`.

```{r}
# Write your code here
```

```{r}
# Write you code here
```

11. Contrast the observed statistic (the real difference in means between `dataset_1` and `dataset_2`) with the values from `perm_diff_means`. Does the permutation test align with the results from the t-test?

```{r}
# Write you answer here
```

