#Install packages and organise sheets to be filled

```{r}
library(tidyr)
library(dplyr)
install.packages("rmarkdown")
# template to be filled
temp<-read.csv("C:/Users/lewis/OneDrive/Documents/MBIO/lewis_temp.csv")
# data from master sheet to be transferred
data<-read.csv("C:/Users/lewis/OneDrive/Documents/MBIO/Lewis_Raw_Sheet.csv")
# Pulling corrected altitude from master database to analysis data sheet
nm <- c("Corrected_height_m",	"Pixel_length")
temp[nm] <- lapply(nm, function(x) data[[x]][match(temp$Filename,data$Filename)])
View(temp)
#write new data back to original temp csv file.
write.csv(temp,"C:/Users/lewis/OneDrive/Documents/MBIO/lewis_temp.csv")
```

#preparing to calculate body volumes for each row, need to clean data first:
```{r}
data<-read.csv("C:/Users/lewis/OneDrive/Documents/MBIO/Lewis_Raw_Sheet.csv")
head(data,20)
dim(data)

#remove rows containg NA value
data <- data %>% drop_na()
dim(data)

# Only keep rows containing S_Alt altitude source 
data<-subset(data, Altitude_Source == 'S_Alt') 
dim(data)

#Remove any measuremetns that didnt pass grading (N)
data<-subset(data, Grade_pass != 'N')
dim(data)

#Check value class and change to numeric
class(data$Corrected_height_m)
data$Corrected_height_m <- as.numeric(data$Corrected_height_m)
class(data$Pixel_length)
```
#Calculate Total length of whale and length of 5% intervals 
```{r}
# Calculate TL
data$TL <- data$Corrected_height_m*data$Pixel_length

#check range to see for anomolies
range(data$TL)

#equation to change 5% intervals from pixel length to absolute length
data<-data %>%  mutate(across(c("Pixel_Width_5","Pixel_Width_10","Pixel_Width_15","Pixel_Width_20","Pixel_Width_25","Pixel_Width_30","Pixel_Width_35","Pixel_Width_40","Pixel_Width_45","Pixel_Width_50","Pixel_Width_55","Pixel_Width_60","Pixel_Width_65","Pixel_Width_70","Pixel_Width_75","Pixel_Width_80","Pixel_Width_85","Pixel_Width_90","Pixel_Width_95"), function(x) x*data$TL/100))
View(data)
```

#isolating 5 % intervals
```{r}
colnames(data)[10:28] <- c("Width.5.proc.m","Width.10.proc.m","Width.15.proc.m","Width.20.proc.m","Width.25.proc.m","Width.30.proc.m","Width.35.proc.m","Width.40.proc.m","Width.45.proc.m","Width.50.proc.m","Width.55.proc.m","Width.60.proc.m","Width.65.proc.m","Width.70.proc.m","Width.75.proc.m","Width.80.proc.m","Width.85.proc.m","Width.90.proc.m","Width.95.proc.m")
```

#Renaming TL and Role.
```{r}
data <- data %>% 
  rename("Total.length.m" = "TL",
         "Rep.class" = "Role")
```
#reorganisng/subsetting and reprganisng columns of interest. changename to BM
```{r}
BM<-as.data.frame(data[,c("Filename","Total.length.m","Rep.class","Width.5.proc.m","Width.10.proc.m","Width.15.proc.m","Width.20.proc.m","Width.25.proc.m","Width.30.proc.m","Width.35.proc.m","Width.40.proc.m","Width.45.proc.m","Width.50.proc.m","Width.55.proc.m","Width.60.proc.m","Width.65.proc.m","Width.70.proc.m","Width.75.proc.m","Width.80.proc.m","Width.85.proc.m","Width.90.proc.m","Width.95.proc.m")])

#confirming widths are numeric
class(BM$Width.5.proc.m)
```


#choosing 5% intervals from 5%-85% of the whale.
```{r}
HW.ratios<-data.frame(
  Measurement.site=seq(5,85,5),
  All.HW.ratio=c(0.89,0.78,0.90,0.97,0.91,0.92,0.93,0.92,0.94,1.00,1.08,1.19,1.34,1.51,1.75,2.03,2.26))

for(i in 1:length(BM[,1])){                                                                         #Runs a loop for every individual in the data frame
  temp.widths<-BM[i,c(which(colnames(BM)=="Width.5.proc.m"):which(colnames(BM)=="Width.85.proc.m"))]  #Extracts the body width measurements for individual i
  temp.heights<-temp.widths*HW.ratios$All.HW.ratio                    #...calculate the height of whale i based on the estimated HW ratios 
  colnames(temp.heights)<-c("Height.5.proc.m","Height.10.proc.m","Height.15.proc.m","Height.20.proc.m","Height.25.proc.m","Height.30.proc.m","Height.35.proc.m","Height.40.proc.m","Height.45.proc.m","Height.50.proc.m","Height.55.proc.m","Height.60.proc.m","Height.65.proc.m","Height.70.proc.m","Height.75.proc.m","Height.80.proc.m","Height.85.proc.m")   #Re-name the column headers for the height measurements
  ifelse(i==1,temp.output<-temp.heights,temp.output<-rbind(temp.output,temp.heights))                 #Extracts the data into a temporary output data frame called "temp.output"
}
BM<-cbind(BM,temp.output)  
```

#Set width and height at 0 and 100%BL from rostrum to be 0
```{r}
BM$Width.0.proc.m<-0
BM$Height.0.proc.m<-0
BM$Width.100.proc.m<-0
BM$Height.100.proc.m<-0
```

#Calculate the body width and height at 90 and 95%BL from rostrum based on linear interpolation between 85 and 100%BL from rostrum
```{r}
BM$Width.90.proc.m<-BM$Width.85.proc.m-(1*(BM$Width.85.proc.m/3))
BM$Width.95.proc.m<-BM$Width.85.proc.m-(2*(BM$Width.85.proc.m/3))
BM$Height.90.proc.m<-BM$Height.85.proc.m-(1*(BM$Height.85.proc.m/3))
BM$Height.95.proc.m<-BM$Height.85.proc.m-(2*(BM$Height.85.proc.m/3))
```

#Re-order data frame (necessary for the next step)
```{r}
BM<-BM[,c(which(colnames(BM)=="Width.0.proc.m"),which(colnames(BM)=="Width.5.proc.m"):which(colnames(BM)=="Width.85.proc.m"),which(colnames(BM)=="Width.90.proc.m"),which(colnames(BM)=="Width.95.proc.m"),which(colnames(BM)=="Width.100.proc.m"),which(colnames(BM)=="Height.0.proc.m"),which(colnames(BM)=="Height.5.proc.m"):which(colnames(BM)=="Height.85.proc.m"),which(colnames(BM)=="Height.90.proc.m"),which(colnames(BM)=="Height.95.proc.m"),which(colnames(BM)=="Height.100.proc.m"),which(colnames(BM)=="Total.length.m"),which(colnames(BM)=="Rep.class"),which(colnames(BM)=="Filename"),which(colnames(BM)=="Year"))] 
Width.col.start<-which(colnames(BM)=="Width.0.proc.m")        #Extracts the column number of the starting width measurement
Height.col.start<-which(colnames(BM)=="Height.0.proc.m")      #Extracts the column number of the starting height measurement
```
#Calculate the body volume of the whales
```{r}
BM$Volume.m3<-NA                            #Creates an empty storage vector for body volume
for(y in 1:length(BM[,1])){                 #Runs a loop for every individual in the data frame
  for(k in 1:(length(seq(0,100,5))-1)){     #Runs a loop for every body segment (volume between two measurement sites), including 0 and 100%
    f.ellipse<-function(x){                   #Formula to calculate the volume of an ellipse        
      (BM[y,Width.col.start+(k-1)]+((BM[y,(Width.col.start+k)]-BM[y,Width.col.start+(k-1)])*x))/2*(BM[y,Height.col.start+(k-1)]+((BM[y,(Height.col.start+k)]-BM[y,Height.col.start+(k-1)])*x))/2*pi
    }
    Volume.temp<-integrate(f.ellipse,lower=0,upper=1)$value*BM$Total.length.m[y]*0.05   #Multiplies the area with the volume of the segment
    ifelse(k==1,Store1<-Volume.temp,Store1<-c(Store1,Volume.temp))                      #Stores the output from each body segment volume estimate
  }
  BM$Volume.m3[y]<-sum(Store1)                                                          #Calculates the body volume from all the body segments for individual i
}
```

# check data
```{r}
plot(BM$Total.length.m,BM$Volume.m3)

#Check data makes sense/that there arent many outliers
data$Volume.m3 <- NA
nm <- c("Volume.m3")
data[nm] <- lapply(nm, function(x) BM[[x]][match(data$Filename,BM$Filename)])
View(data)
plot(data$Total.length.m,data$Volume.m3)
```

```{r}
#write the CSV file
write.csv(data, "C:/Users/lewis/OneDrive/Documents/MBIO/Lewis_Raw_Sheet.csv")
```