---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

data_df <- read.csv("your_data_file.csv")


How many lines does the data frame contain?

```{r}
water_chemistry <- read.csv("water_chemistry.csv")
nrow(water_chemistry)
```

How many columns does the data frame contain?

```{r}
ncol(water_chemistry)
```


How many non-NA values does the column PO43- (µM) contain?

```{r}

```

How many values are greater than 0.10 in column NO3- & NO2- (µM)?

```{r}
sum(water_chemistry$NO3_NO2 > 0.1, na.rm = TRUE)
```

Sort the data by column Date in ascending and on column Temp. (˚C) in descending.

```{r}

```

How many column values are greater than 0.10 in column NO3- & NO2- (µM) and less than 25 in column Temp. (˚C)?

```{r}
sum((water_chemistry$NO3_NO2 > 0.1) & (water_chemistry$Temp < 25), na.rm = TRUE)
```

Remove rows that have NA in column NH4+ & NH3 (µM).

```{r}
is.na(water_chemistry$NH4_NH3) #see which values are NA (=TRUE)
water_chemistry[!(is.na(water_chemistry$NH4_NH3)), ] # ! means "not", so instead of returning the values that are TRUE it will return values that are FALSE

```

What is the average value of column Salinity (PSU)?

```{r}

```

What is the minimum value in column TDP (µM)?

```{r}
```

What is the maximum value in column TDN (µM)?

```{r}
max(water_chemistry$TDN)
```

Create a vector containing values from column Turbidity (NTU).

```{r}
turb <- water_chemistry$Turbidity
head(turb)
```

What is the sum of all the values in the column Chl a (µg/L)?

```{r}
sum(water_chemistry$Chl_a, na.rm = TRUE)
```

Filter rows where pH (unit) is greater than 8.

```{r}
water_chemistry[(water_chemistry$pH > 8), ]
```

Create a new column that is the ratio of TDP (µM) to TDN (µM).

```{r}
```

Count how many unique values are there in Temp. (˚C).

```{r}
```

Extract the first 5 rows and columns 3 to 7 into a new data frame.

```{r}
extract <- water_chemistry[seq(1,5), seq(3,7)]
extract
```

Convert the column Salinity (PSU) to a numeric vector.

```{r}
```

Create a matrix with columns PO43- (µM), NO3- & NO2- (µM) and NH4+ & NH3 (µM).

```{r}
```

What is the median of the column DO (ppm)?

```{r}
```

Replace NA values in column Turbidity (NTU) with the mean of the column.

```{r}
```
How many unique dates are in the dataset?

```{r}
```

Create a new data frame containing only columns related to Nitrogen (NO3- & NO2- (µM), NH4+ & NH3 (µM) and TDN (µM)).

```{r}
```

How many rows have Salinity (PSU) greater than 35?

```{r}
```

What is the standard deviation of column pH (unit)?

```{r}
```

Convert the data frame to a matrix and store it in a variable.
```{r}
```


Find the date associated with the highest Chl a (µg/L) value.

```{r}

```

Find the top 3 lowest Temp. (˚C) values and their corresponding Date.

```{r}
temp_increasing <- water_chemistry[order(water_chemistry$Temp), ]
temp_increasing[seq(1,3), "M_D_Y"]
```



