---
title: "In Class Exercises: Probability Distribution"
output:
  pdf_document: default
  html_document: default
---
```{r}
library(ggplot2)
```


### Understanding Coral Health

0. You go on an expedition to sample 10 fish, each of which can either be sick (due to some parasite) or healthy. What is a possible outcome of this experiment? What is the sample space?

```{r}
#The possible outcome is a ratio of 0-10 fish are sick or healthy. We don't have a probability of infection so we can't make an estimate. Sample space = 10
(H,D,H,H,H,H,H,D,H,H)
(1,2,3,4,5,6,7,8,9,10)
```


1. Your colleagues observe that there are 19 times more healthy fish than diseased ones. What is the probability of observing a healthy fish and a diseased one?

```{r}
#1 out of 20 will be sick. So probability is 0.05
p_sick = 1/20
#19 out of 20 will be healthy. So probability is 0.95
p_healthy = 19/20

```

2. Write code in R to simulate a possible outcome of such an experiment.

```{r}
#sick
rbinom(1, 10, 0.05) #1 experiment,  sample size of 10, with 5% probability of being sick

#healthy
rbinom(1, 10, 0.95)
```

3. What is the probability of observing a site with 5 diseased fish and 5 healthy ones?

```{r}
prob_5_5 <- dbinom(5, 10, 1/20) #what is the probability that 5 fish will be sick out of 10?
prob_5_5
#probability of observing the 5 sick and 5 healthy out of 10 fish is EXTREMELY low, alomst impossible.
```


4. Plot the probability distribution of this experiment.

```{r}
x <- 0:10 #sample space
x
y <- dbinom(x, 10, 1/20) #probability of each potential observation
y

df <- data.frame(x, y)

ggplot(df, aes(x, y))+
  geom_bar(stat="identity")
```

5. Based on the plot, what is the most likely outcome? Compute the event's probability.

```{r}
#Based on the plot, the most likely outcome is 0 sick.
dbinom(0, 10, 1/20)
```

6. Provide an example from your discipline where the binomial distribution applies. Identify the experiment, event, sample space, \( p \), \( n \), etc.

```{r}
# Add your answer here

```

### Understanding Atlantic Cod Length Distribution

1. You select 10 fish from the ocean and measure their lengths. What is a possible set of 10 lengths you should expect?

```{r}
# Solution

```

2. Write code in R to simulate a possible outcome of such an experiment.

    ```{r}
    # Solution
    ```

3. What are the probabilities of observing a fish with lengths 83, 105, and 115?

    ```{r}
    # Solution
    ```

```{r}
```

4. Plot the probability distribution for this experiment.

```{r}
```

```{r}
```

5. Based on the plot, what is the most likely fish size? Compute the event's probability.

    ```{r}
    ```

6. Provide an example from your discipline where the Gaussian distribution applies. Identify the experiment, event, sample space, putative mean, and sigma, etc.

    ```{r}
    ```



### Studying Marine Ecosystems

0. You observe a particular region of the ocean for 30 minutes to count the number of jellyfish passing by. What is a possible outcome of this experiment? What is the sample space?

    ```{r}
    ```

1. Write code in R to simulate a possible outcome of such an experiment.

    ```{r}
    # Solution
    ```

2. Historically, colleagues who have ovserved the same site in the past reported an average of 10 jellyfish observed in a 30-minute interval. What is the 
probability of observing exactly 12 jellyfish?

    ```{r}
    # Solution
    ```


3. What is the probability of observing fewer than 5 jellyfish in a 30-minute interval?

```{r}
    # Solution
```




    ```{r}
    # Solution
    ```

4. Plot the probability distribution for this experiment.



```{r}
```

5. Based on the plot, what is the most likely outcome? Compute the event's probability.

    ```{r}
    # Solution
    ```

6. Provide an example from your discipline where the Poisson distribution applies. Identify the experiment, event, sample space, \( \lambda \), etc.

    ```{r}
    ```

